Analysis of the 2005--2015 Earthquake Sequence in Northern Iran Using the Visibility Graph Method

### 2016-08-21 conclusion.tex
Adding the synthetic data of Telesca 2014 (POne) paper. And general review.
Ricardo's comments added.
finished revising the results section

\section{Conclusion}
We studied the earthquake sequence of three tectonic seismic regions of north Iran. Results manifest the strong relationship between $k-M$ slope and Gutenberg-Richter parameters. The variation of $k-M$ slope and $b-value$ with time in sliding windows have similar behavior and drops down before big earthquakes. Significant reduction of $<T_C>$ value before big earthquake is observed. Higher coefficient of variation of $k-M$ slope in comparison with $b-value$'s CV, suggests the $k-M$ slope as better parameter for studying the magnitude time series due to higher sensitivity to window size and threshold magnitude than $b-value$. In general, $k-M$ and $b-value$ relationship preserves the behavior with variation of the window size and threshold magnitude, and could be an alternative approach for clustering the earthquake sequences based on different tectonic seismic zones.

>>>>

\section{Conclusions}

Pending...

% We studied the earthquake sequence of three tectonic seismic regions of northern Iran. Results manifest the strong relationship between  $k-M$  slope and Gutenberg-Richter parameters. The variations of  $k-M$  slope and  b-value  with time in sliding windows have similar behavior and decline before large earthquakes. A significant reduction of  $<T_c>$  value before large earthquakes was observed. The higher coefficient of variation of  $k-M$  slope in comparison with  b-value  suggests that the  $k-M$  slope is a better parameter for studying the magnitude time series due to higher sensitivity to window size and threshold magnitude than  is found in the $b-value$. Sensitivity analysis revealed that the VG parameters behave independently than threshold magnitude and number of events. In general,  the $k-M$  and  $b-value$  relationship preserves the behavior with the variation in the window size and threshold magnitude, and could be an alternative approach for clustering earthquake sequences based on different tectonic seismic zones.

### 2016-08-21 results.tex
finished revising the results section

According to these results, a universal relationship between the $b$-value and the $k$-$M$ slope ($m$) can be expressed as:
% 
\begin{equation}
	b = 000 + 000 m \, .
	\label{eq:universal.bm}
\end{equation}

\begin{figure}[h]%[t]
	\centering
	\includegraphics[width=0.45\textwidth]{figures/pdf/figure-07} 
	\caption{Relationship between $k-M$ slope and $b-value$ of three Iranian tectonic seismic zones (this study) and three other studies of Mexican zones \citep{Telesca2013}, Pannonia zones \citep{Telesca2014}, and synthetic data \citep{Telesca2014-pone}. The lines represent the linear regression fit of data. The correlation coefficients (R) are represented for different combination of data.}
	\label{fig:regression}
\end{figure}

@@

\begin{figure*}%[t]
	\centering
	\includegraphics[width=0.9\textwidth]{figures/pdf/figure-08} 
	\caption{Relationship between $k-M$ and $b-value$ of three Iranian tectonic seismic regions for 200 random sequence. The numbers on the mean and standard deviation plots are coefficient of variations in percent $(standard \ deviation / mean)$.}
	\label{fig:random}
\end{figure*}

Figure \ref{fig:random}a shows scattered points for all the individual $k$-$M$ slope and $b$-value pairs for all the sub-sequences that were randomly picked from the initial catalogs of each region in northern Iran. Figures \ref{fig:random}b and \ref{fig:random}c show the variability of the $k$-$M$ slope and $b$-value, respectively. These latter figures show the mean values of each parameter for all the random sub-sequences and the amplitudes of $\pm 1$ standard deviation, as well as the coefficients of variation (in percentage). Figure \ref{fig:random}a also includes the data points corresponding to the complete catalogs (solid symbols); the points corresponding to the mean values of the $k$-$M$ slope and $b$, from Figs.~\ref{fig:random}b and \ref{fig:random}c; and the linear regression from equation \ref{eq:universal.bm} as well as the linear regression obtained using the random sub-catalog visual graph analysis for northern Iran, which corresponds to:
% 
\begin{equation}
	b = 000 + 000 m \, .
	\label{eq:iran.bm}
\end{equation}

The results presented in Fig.~\ref{fig:random} indicate there exist only but a small bias (between 5 and 10 percent) depending on the analysis being done with multiple sub-sequences as opposed to the complete catalog sequences. In particular, $k$-$M$ slopes are slightly smaller when done with the random sequence analysis; whereas $b$-values seem more stable. This is also reflected by the differences between the linear regressions for the data of northern Iran exclusively, as opposed to the regression obtained using the data points from this and other studies. In this case, we note that the universal regression increases at a smaller rate than the regression obtain only for northern Iran, as can be seen from the difference between the slope coefficients in equations (\ref{eq:universal.bm}) and (\ref{eq:iran.bm}).

\begin{figure*}%[t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-09} 
	\caption{ Variation of $k-M$ slope  and $b-value$ (Left), and $ < T_c >$ (Right) as a function of time for three tectonic seismic regions in north Iran. Black arrows show the occurrence of the major earthquakes. Numbers on arrows indicate the moment magnitude of the earthquake (see Fig.~\ref{fig:mag-time} for reference).  The magnitude-time series are presented as a filled plot blew each plot. The variation of the heigh of filled plots are corresponding to the event magnitudes. Decreasing the value of  $<T_c>$  ,  $k-M$ , and  $b-value$  before large earthquakes predominantly happened in all tectonic seismic regions.}
	\label{fig:tc}
\end{figure*}

We now investigate the relationship of parameters obtained through the visibility graph analysis and the seismicity of northern Iran with time. Figure \ref{fig:tc} shows...

% =========

% the relationship between  $k-M$  slope and  $b-value$  of the randomly selected data and the statistical parameters for variables. Although changing the number of events and the threshold magnitude is slightly changing the results, however,  the results are fairly well clustered for each tectonic seismic region.

% =========

% We compared the $k$-$M$ slope with the $b$-value of the Gutenberg-Richter law through the entire catalog and in the sliding windows in time.

% Comparing the result with the Guerrero region with reduced random sequence,  \citet{Telesca2013}  found that the number of sequences did not affect the $k-M$ slope.

% Fig.~\ref{fig:random} shows the relationship between  $k-M$  slope and  $b-value$  of the randomly selected data and the statistical parameters for variables. Although changing the number of events and the threshold magnitude is slightly changing the results, however,  the results are fairly well clustered for each tectonic seismic region.

% The coefficient of variation of parameters is shown in Fig.~\ref{fig:random}. The  $k-M$  slope shows the higher coefficient of variation which indicates the fact that the  $k-M$  slope is more dependent on the window size and threshold magnitude than on the a-  and  b-values , which suggests that  the $k-M$  slope can better represent the dynamic characteristics of the magnitude-time series.

% *******

% Fig.\ref{fig:tc} shows the variation of the  $k-M$  slope and  $b-value$  with time. Due to lower number of events in Azerbaijan's catalog, we choose 20 events as the window length with a shift of one event between two successive windows. With higher window length we loose the variation of considerable amount of information in time (e.g. with n=40 we loose almost half of the Azerbaijan catalog). Due to taking average values of successive windows, very small window length can not display the seismicity behavior especially before big earthquake. Maintaining consistency, we use the window length equal to 20 events for all three regions. The calculated parameters of each window are associated with the time of occurrence of the last event in the sliding window. The  $k-M$  slope and  $b-value$  in all tectonic seismic regions are very similar. In general, in all regions the $b-value$  and  $k-M$  slope drop considerably before large earthquakes. The decline in the  $b-value$  before large earthquakes has been studied in many regions \citep{Wyss2000, Wyss2006, Schorlemmer2005, Chan2012}.

% \citet{Telesca2016}  observed the decrease in  $<T_c>$  before the large earthquake of the western India earthquake sequence. Fig.\ref{fig:tc}  shows the time variation of  $<T_c>$   in the study area. The reduction of  $<T_c>$   before large earthquakes is clearly visible.  \citet{Telesca2016}  demonstrated that the decrement of  $<T_c>$   before a large earthquake is independent of window size and threshold magnitude. Decreasing the value of  $<T_c>$  ,  $k-M$ , and  $b-value$  before large earthquakes predominantly happened in all tectonic seismic regions (see Fig.\ref{fig:tc} for reference.)

>>>>

According to these results, a universal relationship between the $b$-value and the $k$-$M$ slope ($m$) can be expressed as:
% 
\begin{equation}
	b = 0.078 + 0.085 m \, .
	\label{eq:universal.bm}
\end{equation}

\begin{figure}[h]%[t]
	\centering
	\includegraphics[width=0.45\textwidth]{figures/pdf/figure-07} 
	\caption{Correlation between $k$-$M$ slope and $b$-value as drawn from the results of the present study for the region of northern Iranian and three other previous studies, including analysis of the Mexican subduction zone \citep{Telesca2013}, the Pannonian seismic zone \citep{Telesca2014}, and results from two experiments \citep{Telesca2014-pone}. The lines represent linear regressions obtained to fit the different data points, considering different combinations. The values of the correlation coefficient, $R$, are indicated for each regression line. The color version of this figure is available only in the electronic edition.}
	\label{fig:regression}
\end{figure}

@@

\begin{figure*}%[t]
	\centering
	\includegraphics[width=0.9\textwidth]{figures/pdf/figure-08} 
	\caption{\textbf{(a)} Correlation between $k$-$M$ slope and $b$-values for 200 random sequences extracted from the initial catalogs of the three northern Iranian seismic regions considered in this study (scattered symbols), including the mean values (empty symbols with thick border) and the data points obtained from the analysis of the complete catalog (solid symbols), along with the linear regressions of each sample as indicated in the legend. \textbf{(b)} Mean, $\pm1$ standard deviation, and coefficient of variation (in percent) for the $k$-$M$ slope values of the three seismic zones in northern Iran. \textbf{(c)} Same as part (b), but corresponding to the $b$-value. The color version of this figure is available only in the electronic edition.}
	\label{fig:random}
\end{figure*}

Figure \ref{fig:random}a shows scattered points for all the individual $k$-$M$ slope and $b$-value pairs for all the sub-sequences that were randomly picked from the initial catalogs of each region in northern Iran. Figures \ref{fig:random}b and \ref{fig:random}c show the variability of the $k$-$M$ slope and $b$-value, respectively. These latter figures show the mean values of each parameter for all the random sub-sequences and the amplitudes of $\pm 1$ standard deviation, as well as the coefficients of variation (in percentage). Figure \ref{fig:random}a also includes the data points corresponding to the complete catalogs (solid symbols); the points corresponding to the mean values of the $k$-$M$ slope and $b$, from Figs.~\ref{fig:random}b and \ref{fig:random}c; and the universal linear regression from equation (\ref{eq:universal.bm}) (thick line), as well as the linear regressions for northern Iran when using the complete catalogs (dashed, thin line) and the random sub-catalogs (continuous, thin line).

The comparison between the mean values of the random sequences analysis versus the complete catalogs presented in Fig.~\ref{fig:random} indicate there exist only a small bias, which is well within the standard deviation of the different value samples. We note that the values of the $k$-$M$ slope are slightly smaller when done with the random sequence analysis; whereas $b$-values seem more stable. The comparisons of the regressions, however, seem to indicate that the analysis of random sub-catalogs leads to a result more in line with the universal results obtained when considering multiple seismic zones. The linear regress of the random sub-catalogs yields the following equation:
% 
\begin{equation}
	b = 0.080 + 0.093 m \, .
	\label{eq:iran.bm}
\end{equation}

Note that equations (\ref{eq:universal.bm}) and (\ref{eq:iran.bm}) have share similar intercepts with the $b$-value axis, and only slightly different slopes, with the analysis for northern Iran leading to the $b$-value to increase more rapidly with $m$ than the universal data regression suggests. The similarity, however, is a positive sign of the stability of the method.

\begin{figure*}%[t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-09} 
	\caption{Variation of <$T_c$>, $k$-$M$ slope and $b$-value as with respect to time for three seismic zones for a moving window analysis of the visibility graphs of sub-sequences of 20 consecutive events, along with the magnitude of events in the 2005--2015 period. Black triangles indicate the occurrence of major earthquakes in each of the regions, with the corresponding magnitude at the top of each symbol. In the middle frame of each region, the dashed line corresponds to the $k$-$M$ values, whereas the continuous plot shows the variation of the $b$-value. The color version of this figure is available only in the electronic edition.}
	\label{fig:tc}
\end{figure*}

We now investigate the relationship of the parameters obtained through the visibility graph analysis with time. Here, the visibility graph analysis is done by windows of equal number of events moving in time across the catalog sequence. The number of events in the window is kept fixed independently of the time between them, and the results of the visibility graph analysis are associated with the last event in the window. In this case, we are interested in using a small number of events to capture the relevance of each new event as the window moves in time. We tried different numbers and finally chose 20 events as the moving window sequence size. This selection was also convenient given the smaller total number of declustered events in the catalog of Azerbaijan. When we tested larger number of events, we found that we could not properly observe the changes in the different parameters with time. On the other hand, when we used smaller sequence sizes, the values deviated more significantly from the random and complete catalog results obtained previously. At this point it is also of interest to compute the value of <$T_c$> explained int he Methodology section.

Figure \ref{fig:tc} shows the variation of <$T_c$>, the $k$-$M$ slope, and the $b$-value with time for each of the three seismic zones in northern Iran. Note that the behavior of the $k$-$M$ slope and the $b$-value is very similar along time for all three zones. This is consistent with the results discussed above. In the figure we also highlight the occurrence of relevant earthquakes in each seismic zone in the last decade. We note that there seems to be a correlation between the behavior of the $b$-value in time with the occurrence of some of these events. In particular, some of the events in the figure seem to coincide with a drop in the $b$-value.

The decline in $b$-value before large earthquakes has been studied in other regions before \citep[e.g.,][]{Wyss2000, Wyss2006, Schorlemmer2005, Chan2012}. In the context of the visibility graph analysis, \citet{Telesca2016} observed a decrease in <$T_c$> before the large earthquake of the western India earthquake sequence. We recognize, however, that the lack of larger events in our region of interest in the last decade prevents us from drawing a stronger conclusion on this regard.

### 2016-08-21 title-and-authors.tex
improved abstract
almost done with seismicity section
finished revising the results section

\pretitle{
    \vspace{-30pt}
    \begin{flushleft}
    \usefont{T1}{phv}{bc}{it}
    \Huge
}

\posttitle{
    \end{flushleft}
}

\preauthor{
    \begin{flushleft}
    \vspace{3pt}
    \usefont{T1}{phv}{bc}{n}
    \LARGE
}

\postauthor{
    \end{flushleft}
    \vspace{0.3em}
    \section{Abstract}
    \vspace{-3em}
}

\predate{
    \begin{flushleft}
    \makeatletter
}

\postdate{
    \end{flushleft}
}

\title{%
    Analysis of 2005--2015 Earthquake Sequence in Northern Iran Using Visibility Graph Approach
}

\makeatletter
\if@twocolumn
    \date{}
    \renewcommand*\@fnsymbol[1]{\the#1}
    \author{by Naeem Khoshnevis, Shima Azizzadeh-Roodpish, Ricardo Taborda, and Luciano Telesca}
\else
    \author{
    	Naeem Khoshnevis,\myaffil{1,$\ast$} Shima Azizzadeh-Roodpish,\myaffil{1,2} Ricardo Taborda\myaffil{1,2} and Luciano Telesca\myaffil{3}\\
    	\normalsize\normalfont    	    	
        \vspace{20pt}
    	1.~
    	Center for Earthquake Research and Information\\
    	\hspace{1.1em} The University of Memphis, Memphis, TN 38152, U.S.A.\\
        \vspace{20pt}
        2.~
        Department of Civil Engineering\\
        \hspace{1.1em} The University of Memphis, Memphis, TN 38152, U.S.A.\\
        \vspace{20pt}
        3.~
        Institute of Methodologies for Environmental Analysis\\
        \hspace{1.1em} National Research Council, Tito, Italy\\
        \vspace{20pt}
        *.~
    	Corresponding author\\
    	\hspace{1.1em} Email: nkhshnvs@memphis.edu\\
        \vspace{40pt}
    }
\fi
\makeatother

>>>>

\title{%
    Analysis of the 2005--2015 Earthquake Sequence in Northern Iran Using the Visibility Graph Method
}

\makeatletter
\if@twocolumn
    \date{}
    \renewcommand*\@fnsymbol[1]{\the#1}
    \author{by Naeem Khoshnevis, Ricardo Taborda, Shima Azizzadeh-Roodpish, and Luciano Telesca}
\else
    \author{
        Naeem Khoshnevis,\myaffil{1,$\ast$} Ricardo Taborda\myaffil{1,2} Shima Azizzadeh-Roodpish,\myaffil{1,2} and Luciano Telesca\myaffil{3}\\
        \normalsize\normalfont    	    	
        \vspace{20pt}
        1.~
        Center for Earthquake Research and Information\\
        \hspace{1.1em} The University of Memphis, Memphis, TN 38152, U.S.A.\\
        \vspace{20pt}
        2.~
        Department of Civil Engineering\\
        \hspace{1.1em} The University of Memphis, Memphis, TN 38152, U.S.A.\\
        \vspace{20pt}
        3.~
        Institute of Methodologies for Environmental Analysis\\
        \hspace{1.1em} National Research Council, Tito, Italy\\
        \vspace{20pt}
        *.~
        Corresponding author\\
        \hspace{1.1em} Email: nkhshnvs@memphis.edu\\
        \vspace{40pt}
    }
\fi
\makeatother
